---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Peer Assessment 1
=======================
###Loading and preprocessing the data
```{r}
data<-read.csv("activity.csv")
data$date <-as.Date(data$date,"%Y-%m-%d")
```
###What is mean total number of steps taken per day?
* Calculate the total number of steps taken per day
```{r,results='hide'}
library(dplyr)
steps_per_day <-filter(data,!is.na(steps)) %>%group_by(date) %>% summarise(TotalSteps=sum(steps))
```
* Histogram of the total number of steps taken each day
```{r,fig.height=5}
hist(steps_per_day$TotalSteps,main=" ",xlab="Total Number of Steps Taken Daily",col="Blue")
```

* Mean of total number of steps taken each day
```{r}
mean(steps_per_day$TotalSteps);
```
* Median total number of steps taken each day
```{r}
median(steps_per_day$TotalSteps);
```
###What is the average daily activity pattern?
* A time series plotof the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
avg_steps<-filter(data,!is.na(steps)) %>%group_by(interval) %>% summarise(AvgSteps=mean(steps))
```
* The plot
```{r,fig.height= 5}
plot(avg_steps$AvgSteps~avg_steps$interval,type='l',xlab="5-Minute Interval",ylab="Average Number of Steps Count",main="Average Number of Steps Taken Averaged Across All Days",col="red")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avg_steps[which.max(avg_steps$AvgSteps), ]
```

###Imputing missing values
* Calculate and report the total number of missing values in the dataset
```{r}
data_NA <- sum(is.na(data))
data_NA
```
* Devise a strategy for filling in all of the missing values in the dataset. I will use the mean for that 5-minute interval.
```{r}
data <- merge(data,avg_steps, by = "interval")
nas <- is.na(data$steps)
data$steps[nas] <- data$AvgSteps[nas]
data_filled <- data[, c(1:3)]
```

* Calculate the total number of steps taken per day after Imputating NAs.
```{r}
steps_per_day_filled <-data_filled %>%group_by(date) %>% summarise(TotalSteps=sum(steps))
```
* A histogram of the total number of steps taken each day after Imputating NAs.
```{r,fig.height= 5}
hist(steps_per_day_filled$TotalSteps,main=" ",xlab="Total Number of Steps Taken Daily after After Imputating NAs",col="Blue")
```

* Mean of total number of steps taken each day for new data set
```{r}
mean(steps_per_day_filled$TotalSteps);
```
* Median total number of steps taken each day for new data set
```{r}
median(steps_per_day_filled$TotalSteps);
```
* What is the impact of imputing missing data on the estimates of the total daily number of steps?
Answer: Nothing, because we substitute the missing values with the mean values.

###Are there differences in activity patterns between weekdays and weekends?
* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data_filled$weektime <- as.factor(ifelse(weekdays(data_filled$date) %in% c("Saturday","Sunday"),"weekend", "weekday"))
```

* A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r, fig.height= 5}
library("lattice")
avg_steps_new<-data_filled %>%group_by(interval,weektime) %>% summarise(AvgSteps=mean(steps))
xyplot(AvgSteps ~ interval | weektime, avg_steps_new, type = "l", layout = c(1, 2), xlab = "5-minute Interval", ylab = "Average number of steps")
```
